---
title: "class_6_coding_v1"
author: "Thomas Zwiller"
date: "2024-02-02"
output: html_document
---

**1. Import iGraph, SNA, gggraph**
```{r message=FALSE, warning=FALSE}
#Loading the packages I need
library(sna)
library(igraph) 
library(ggraph)
library(gplots)
```

**2. Import Nodes and Edge Data**
```{r}
#Loading the nodes and edges I need
nodes <- read.csv("/Users/TomTheIntern/Desktop/Mendoza/Mod 4/Networks/Lab 2/nodelist.csv")
summary(nodes)
edges <- read.csv('/Users/TomTheIntern/Desktop/Mendoza/Mod 4/Networks/Lab 2/edgelist.csv')
summary(edges)
```
**3. Create Directed iGraph and NetSNA**

```{r}
#Making the iGraph directed
net <- graph_from_data_frame(edges, directed = T, vertices = nodes)
net

#Making the iGraph un-directed
un_net <- graph_from_data_frame(edges, directed = F, vertices = nodes)
un_net
s_un_net <- simplify(un_net)
s_un_net
is_simple(s_un_net)

#making the NetSNA
net_sna <- network(edges, matrix.type = "edgelist", directed = T)
net_sna
```
**4. Graph the Network**

```{r}
#graphing the initial network
ggraph(net) + 
  geom_edge_link0(color = "black", alpha = .5) +
  geom_node_point(fill = "lightblue", color = "darkblue", shape = 21, 
                  size = igraph::degree(net, mode = "all"))
```
<br>

**5. Make 3 Community Detection Algorithms**
<br>
*Spin-Glass*
```{r}
## Spinglass####
?cluster_spinglass
#^This function has a gamma parameter that we can tweak!!
set.seed(1)
csg <- cluster_spinglass(net)
modularity(csg) 
membership(csg) 
length(csg)
sizes(csg)
```

```{r}
# plot####
colors <- colorpanel(length(csg), low = "#efd9d3", high = "#57424a")
ggraph(net) + 
  geom_edge_link0(color = "black", alpha = .5) +
  geom_node_point(aes(fill = as.factor(membership(csg))),
                  size = igraph::degree(net, mode = "all"), 
                  color = "black", shape = 21) +
  scale_fill_manual(values = c(colors)) +
  ggnetwork::theme_blank() +
  theme(legend.position = "none")
```

<br>
*Louvain Method*
```{r}
## Louvain method####
?cluster_louvain 
# has resolution parameter we can tweak!
set.seed(1)
clv <- cluster_louvain(un_net)
modularity(clv) 
membership(clv) 
communities(clv)
length(clv)
sizes(clv)
```

```{r}
# plot####
colors <- colorpanel(length(clv), low = "#efd9d3", high = "#57424a")
ggraph(net) + 
  geom_edge_link0(color = "black", alpha = .5) +
  geom_node_point(aes(fill = as.factor(membership(clv))),
                  size = igraph::degree(net, mode = "all"), 
                  color = "black", shape = 21) +
  scale_fill_manual(values = c(colors)) +
  ggnetwork::theme_blank() +
  theme(legend.position = "none")
```

```{r}
# tuning parameters####
set.seed(1)
clv_1 <- cluster_louvain(un_net, resolution = 0.25)
length(clv_1)
clv_2 <- cluster_louvain(un_net, resolution = 0.55)
length(clv_2)
clv_3 <- cluster_louvain(un_net, resolution = 0.75)
length(clv_3)
clv_4 <- cluster_louvain(un_net, resolution = 1)
length(clv_4)
clv_5 <- cluster_louvain(un_net, resolution = 1.25)
length(clv_5)
clv_6 <- cluster_louvain(un_net, resolution = 1.5)
length(clv_6)

colors <- colorpanel(length(clv_1), low = "#efd9d3", high = "#57424a")
clv_1_plot <- ggraph(un_net) +
  geom_edge_link0(color = "black", alpha = .5) +
  geom_node_point(aes(fill = as.factor(membership(clv_1))),
                  size = igraph::degree(un_net, mode = "all"), 
                  color = "black", shape = 21) +
  scale_fill_manual(values = c(colors)) +
  ggnetwork::theme_blank() +
  geom_node_text(aes(label = as.factor(membership(clv_1))), repel = T) +
  theme(legend.position = "none")

colors <- colorpanel(length(clv_2), low = "#efd9d3", high = "#57424a")
clv_2_plot <- ggraph(un_net) +
  geom_edge_link0(color = "black", alpha = .5) +
  geom_node_point(aes(fill = as.factor(membership(clv_2))),
                  size = igraph::degree(un_net, mode = "all"), 
                  color = "black", shape = 21) +
  scale_fill_manual(values = c(colors)) +
  ggnetwork::theme_blank() +
  geom_node_text(aes(label = as.factor(membership(clv_2))), repel = T) +
  theme(legend.position = "none")

colors <- colorpanel(length(clv_3), low = "#efd9d3", high = "#57424a")
clv_3_plot <- ggraph(un_net) +
  geom_edge_link0(color = "black", alpha = .5) +
  geom_node_point(aes(fill = as.factor(membership(clv_3))),
                  size = igraph::degree(un_net, mode = "all"), 
                  color = "black", shape = 21) +
  scale_fill_manual(values = c(colors)) +
  ggnetwork::theme_blank() +
  geom_node_text(aes(label = as.factor(membership(clv_3))), repel = T) +
  theme(legend.position = "none")

colors <- colorpanel(length(clv_4), low = "#efd9d3", high = "#57424a")
clv_4_plot <- ggraph(un_net) +
  geom_edge_link0(color = "black", alpha = .5) +
  geom_node_point(aes(fill = as.factor(membership(clv_4))),
                  size = igraph::degree(un_net, mode = "all"), 
                  color = "black", shape = 21) +
  scale_fill_manual(values = c(colors)) +
  ggnetwork::theme_blank() +
  geom_node_text(aes(label = as.factor(membership(clv_4))), repel = T) +
  theme(legend.position = "none")

colors <- colorpanel(length(clv_5), low = "#efd9d3", high = "#57424a")
clv_5_plot <- ggraph(un_net) +
  geom_edge_link0(color = "black", alpha = .5) +
  geom_node_point(aes(fill = as.factor(membership(clv_5))),
                  size = igraph::degree(un_net, mode = "all"), 
                  color = "black", shape = 21) +
  scale_fill_manual(values = c(colors)) +
  ggnetwork::theme_blank() +
  geom_node_text(aes(label = as.factor(membership(clv_5))), repel = T) +
  theme(legend.position = "none")

colors <- colorpanel(length(clv_6), low = "#efd9d3", high = "#57424a")
clv_6_plot <- ggraph(un_net) + 
  geom_edge_link0(color = "black", alpha = .5) +
  geom_node_point(aes(fill = as.factor(membership(clv_6))),
                  size = igraph::degree(un_net, mode = "all"), 
                  color = "black", shape = 21) +
  scale_fill_manual(values = c(colors)) +
  ggnetwork::theme_blank() +
  geom_node_text(aes(label = as.factor(membership(clv_6))), repel = T) +
  theme(legend.position = "none")

library(ggpubr)
ggarrange(clv_1_plot, clv_2_plot, clv_3_plot, clv_4_plot, clv_5_plot, clv_6_plot,
          nrow = 2, ncol = 3)
```
<br>
*Info Map*
```{r}
## InfoMAP####
?cluster_infomap
set.seed(1)
cim <- cluster_infomap(net)
modularity(cim) 
membership(cim) 
length(cim)
sizes(cim)
```

```{r}
# plot####
colors <- colorpanel(length(cim), low = "#efd9d3", high = "#57424a")
ggraph(net) + 
  geom_edge_link0(color = "black", alpha = .5) +
  geom_node_point(aes(fill = as.factor(membership(cim))),
                  size = igraph::degree(net, mode = "all"), 
                  color = "black", shape = 21) +
  scale_fill_manual(values = c(colors)) +
  ggnetwork::theme_blank() +
  theme(legend.position = "none")
```
<br>
**7. Best Fitting Algorithm**

I started by removing the Info-Map from contention as it only divided the network into two communities when there should have been three communities. 

Originally, I ended up settling on Spin-Glass because it seemed to be the best at deciding where the central node (myself) should go. 

However, the Louvian method scored a 0.39125 Modularity, which was the best of the three, and was my final selection. 

<br>
**8. Plot the Final**

```{r}
colors <- colorpanel(length(clv_6), low = "lightblue", high = "darkblue")
clv_6_plot <- ggraph(un_net) + 
  geom_edge_link0(color = "black", alpha = .5) +
  geom_node_point(aes(fill = as.factor(membership(clv_6))),
                  size = igraph::degree(un_net, mode = "all"), 
                  color = "black", shape = 21) +
  scale_fill_manual(values = c(colors)) +
  ggnetwork::theme_blank() +
  geom_node_text(aes(label = as.factor(membership(clv_6))), repel = T) +
  theme(legend.position = "none")

library(ggpubr)
ggarrange(clv_6_plot, nrow = 1, ncol = 1)
```




<br>
<br>
<br>
<br>
<br>
<br>
<br>