---
title: "Class 7 Lab"
author: "Thomas Zwiller"
date: "2024-04-08"
output: html_document
---

**#1. Install lme4 and lmerTest packages**
```{r message=FALSE, warning=FALSE}
#importing the libraries I need
library(lme4)
library(lmerTest)
library(gplots)
library(dplyr)
library(igraph)
library(performance)
```
<br>

**#2. Importing Nodes and Edges**
```{r}
#reading in the nodes data

airport_data <- read.csv('/Users/TomTheIntern/Desktop/US_Airport_500.csv')

nodes <- read.csv('/Users/TomTheIntern/Desktop/Mendoza/Mod 4/Networks/Lab 2/nodelist.csv')
summary(nodes)

#reading in the edges
edges <- read.csv('/Users/TomTheIntern/Desktop/Mendoza/Mod 4/Networks/Lab 2/edgelist.csv')
summary(edges)
```
**#3. Merge the edge and node data frames**
```{r}
#merging the nodes and edges data by ego id
edges_nodes <- merge(edges, nodes, by.x  = 'ego_num', by.y = 'ID')
```
<br>

**#4. Random Intercept Model**
```{r}
### 1. Random intercept model where I try to predict strencth with the relationship type
model1 <- lmer(strength ~ type + (1 | ego_num), #
               data = edges_nodes, REML = T, 
               na.action = na.omit)
summary(model1)
```
<br>
**#5. Interpret the Model Results**

Because the default type is friend, if you are a friend there is a slight dip in the relationship strength compared to the family type. However, it is not a statistically significant predictor, and has a relatively small impact on the regression, suggesting that the difference in type is not predictive of the strength of the relationship between egos.

<br>

**#6. Calculate the ICC**
```{r}
#checking the ICC
performance::icc(model1)
```

**#7. Interpret the ICC**
An ICC of 0.096 is rather low, meaning that the relationships within each ego are rather different, which makes sense as I tried to estimate relationships between employee and employer, which would have a different strength. The same is true of Marcus Freeman, who the staff spends a lot of time talking/thinking about and asking questions/observing, but Freeman only really interacts with them at a press conference. This type of relationship also works with fans.


<br>
<br>
<br>
<br>
<br>
<br>
<br>
