---
title: "Class 5 Lab"
author: "Tom Zwiller"
date: "2025-03-30"
output:
  pdf_document: default
  html_document: default
---

**1. Load the SNA and iGraph Packages**

```{r message=FALSE, warning=FALSE}
#Loading the packages I'll need
library(ggpubr)
library(ggraph)
library(dplyr)
library(igraph) 
library(sna)
library(intergraph) 
```
<br>

**2. Load your directed network data**
```{r}
#Importing the node and edge csv's
nodes <- read.csv('/Users/TomTheIntern/Desktop/Mendoza/Mod 4/Networks/Lab 2/nodelist.csv')
summary(nodes)
edges <- read.csv('/Users/TomTheIntern/Desktop/Mendoza/Mod 4/Networks/Lab 2/edgelist.csv')
summary(edges)
```
<br>

**3. Create Directed SNA and iGraph**
```{r}
# Create igraph and sna object and visualize the network
# first list the edgelist, whether the network is directed, and the vertex attributes.
net <- graph_from_data_frame(edges, directed = T, vertices = nodes)

#making the igraph object into an sna object
net_s <- igraph::simplify(net)
net_s_sna <- asNetwork(net_s) 
```
<br>

**4. Calculating Triad Census**

```{r}
#Making the directed Triad Census
#using sna triad for the better labels
sna::triad.census(net_s_sna)
#getting the highest type of triad
colnames(sna::triad.census(net_s_sna))[which.max(sna::triad.census(net_s_sna))]
```
The 102 triad (1 mutual, 0 asymetirc and 2 nulls) was the most prevalent in the network with a total of 113 instances. 

<br>

**5. Age Assortativity of the Network**

```{r}
assortativity(net, V(net)$Age, directed = T)
```
The age assortativity was 0.04203086, meaning I had some older ages associating with other older aged people.

<br>

**6. Conditional Uniform Graph of Transitivity vs Density**

```{r}
#Test transitivity vs density
Cug_Edges <- sna::cug.test(net_s_sna, gtrans, cmode="edges")

#making the plot
plot(Cug_Edges)

#our network
Cug_Edges$obs.stat 

# simulated networks
summary(Cug_Edges$rep.stat) 
```
Compared to the 1000 simulations, my network had a much, much higher density than the average simulation and was still more dense than some of the outlier simulations, meaning my network was much more dense than any of the simulations. 


<br>
<br>
<br>
<br>
<br>
<br>
