---
title: "Class Eight Lab"
author: "Tom Zwiller"
date: "2024-04-10"
output:
  pdf_document: default
  html_document: default
---
**#1. Installing and Loading in the Ergm Package**
```{r message=FALSE, warning=FALSE}
#Installing the libraries we need
library(ergm)
library(dplyr)
library(magrittr)
library(ggraph)
library(sna)
library(btergm)
```
<br>

**#2. Loading in the Node and Edge Data**
```{r}
#Node Data
nodes <- read.csv('/Users/TomTheIntern/Desktop/Mendoza/Mod 4/Networks/Lab 2/nodelist.csv')
summary(nodes)
#Edge Data
edges <- read.csv('/Users/TomTheIntern/Desktop/Mendoza/Mod 4/Networks/Lab 2/edgelist.csv')
summary(edges)

net_sna <- network(edges, matrix.type = "edgelist", 
                   directed = T, vertex.attr = nodes)
net_sna
```
<br>

**#3. Adding Two Terms to the Model**
```{r message=FALSE, warning=FALSE}
#basic ERGM
model1 <- ergm(net_sna ~ edges +  nodematch("Gender") + gwesp(0.05, fixed = TRUE))
```
<br>

**#4. Interpret the Model**
```{r}
summary(model1)
```

*Goodness of Fit*
```{r}
gof.model1_btergm <- btergm::gof(model1, nsim = 100, # checking goodness of fit of model estimates against 100 simulated networks
                                   statistics = c(deg, odeg, ideg, triad.directed))
gof.model1_btergm 
```

I tinkered with the model for a bit, swapping variables until I found a paring that was statistically significant and finally found that nodematch worked well with gwesp.

Nodematch focused on Gender makes the most sense; I tended to have more men in my network and those men generally had relationships with other men; the only exceptions were myself and my girlfriend, her father and her mother and my friend and his girlfriend. 

And then Gwesp was strong because most of the time two nodes had a handful of shared partners, hence it being statistically significant. 

All things considered, the model had a low AIC (134) and BIC (142.7), and was relatively accurate

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>