---
title: "class_4_lab"
author: "Thomas Zwiller"
date: "2025-03-26"
output: html_document
---

**1. Loading the igraph package**
```{r}
#Loading the libraries I need 
library(igraph) 
library(ggraph)
```

**2. Loading in the ego-centric data**
```{r}
#Importing the node and edge csvs
nodes <- read.csv("/Users/TomTheIntern/Desktop/Mendoza/Mod 4/Networks/Lab 2/nodelist.csv")
summary(nodes)
edges <- read.csv('/Users/TomTheIntern/Desktop/Mendoza/Mod 4/Networks/Lab 2/edgelist.csv')
summary(edges)
```
**3 Creating the igraph object**

```{r}
#making the directed igraph
net <- graph_from_data_frame(edges, directed = T, vertices = nodes)
net

#making the un-directed igraph
un_net <- graph_from_data_frame(edges, directed = F, vertices = nodes)
un_net
```

**4. Choose two centrality measures and apply them to your network**
I am using Degree and Eigenvector as my two centrality measures

**5. Visualize your network (one graph for each measure) with node size representing your chosen measures**

**Degree**
```{r}
ggraph(net) + 
  geom_edge_link0(color = "black", alpha = .5) +
  geom_node_point(fill = "lightblue", color = "darkblue", shape = 21, 
                  size = igraph::degree(net, mode = "all"))
```

**Eigenvector**
```{r}
ggraph(un_net) + 
  geom_edge_link(color = "black", alpha = .5) +
  geom_node_point(fill = "lightblue", color = "darkblue", shape = 21, 
                  size = igraph::eigen_centrality(un_net)$vector*16)
```

**6. Describe what the measure tells you about the people in your network**
<br>

**What Degree tells me about the people in my network**
Unsurprisingly, I had the most degrees, which made my node the biggest. Generally, the different communities were relatively the same size: 

the football network (my co-workers) were all consistent because they were connected. 
my girlfriend's family generally tended to be smaller, as they only had 2-3 connections compared to football having 4-5

my res-life coworkers were the smallest, with my co-workers girlfriend having just one connection, making it the smallest, and about as close to an isolet as possible.

<br>
**What Eigenvector tells me about the people in my network**

The Eigenvector measure looks relatively similar to the degree measure, though there are a few key differences. My girlfriend has a much bigger circle in the nodes, but because most of her connections were to other relatively un-connected nodes, she had a smaller representation. 

Comparatively, the football network was actually bigger because they all knew me, and each other, which helped improve their score. 

My res-life co-workers were relatively smaller than their node score, however my co-worker who works with the ND bad (and would theoretically know Marcus Freeman) had a bit of a higher score.

<br>
<br>
<br>
<br>
<br>
<br>
<br>
