---
title: "ITAO_80870_class_10_canvas"
author: "T. Zwiller"
date: "2024-04-17"
output:
  pdf_document: default
  html_document: default
---

**#1. Loading iGraph, GGraph, and NetDiffuse**
```{r message=FALSE, warning=FALSE}
#loading in the needed libraries
library(igraph)
library(ggraph)
library(netdiffuseR)
```
<br>

**#2. Loading in Egocentric Network Data**
```{r}
#loading in the edge and node data
edges <- read.csv('/Users/TomTheIntern/Desktop/Mendoza/Mod 4/Networks/Lab 2/edgelist.csv')
nodes <- read.csv('/Users/TomTheIntern/Desktop/Mendoza/Mod 4/Networks/Lab 2/nodelist.csv')
```
<br>

**#3. Making a directed iGraph Object**
```{r}
#making the iGraph
net <- graph_from_data_frame(edges, directed = F, vertices = nodes)
net
```
<br>

**#4. Visualize Your Network**
```{r}
#making the plot
ggraph(net) +
  geom_edge_link(color = "black", alpha = 0.3) +
  geom_node_point(color = "black", fill = "#efd9d3", shape = 21) +
  theme_graph()
```
<br>

**#5. Add 4 Ties, 4 Nodes, Remove 3 Ties, 3 Nodes And Rewire Network**
```{r}
library(igraph)

set.seed(000027)

# 1. Add your new ties
net2 <- igraph::add_edges(net, c(1, 8, 2, 9, 3, 12, 7, 12))

# 2. Delete those edges by ID
net2 <- igraph::delete_edges(net2, c(1, 10, 3, 4, 10, 11))

# 3. Add 4 brand-new vertices (scalar count)
net2 <- igraph::add_vertices(net2, 4)

# 4. Remove vertices 12, 5, and 4 by position
net2 <- igraph::delete_vertices(net2, c(12, 5, 4))

# 5. Re-name each node sequentially so name index stay in sync
V(net2)$name <- as.character(seq_len(vcount(net2)))

# 6. Randomly rewire 50% of edges, no loops
net2 <- igraph::rewire(net2, each_edge(p = 0.5, loops = FALSE))
```
<br>

**#6. Plot the new network with GGraph**
```{r}
#plotting the new graph
ggraph(net2) +
  geom_edge_link(color = "black", alpha = 0.3) +
  geom_node_point(color = "black", fill = "#efd9d3", shape = 21) +
  theme_graph()
```
<br>

**#7. Run a threshold model over the original vs new graph**
```{r}
#converting the og plot into a matrix
net_matrix <- as_adjacency_matrix(net)
n_nodes <- nrow(net_matrix) 

#converting the new plot into a matrix
net2_matrix <- as_adjacency_matrix(net2)
n2_nodes <- nrow(net2_matrix) 
```
<br>

Graphing Original Network
```{r}
#setting the seed
set.seed(1)

#running the diffusion threshold model
diffnet_il <- rdiffnet(
  t = 10, 
  seed.graph = net_matrix,
  seed.nodes = c(1, 7, 11), # chosen nodes
  seed.p.adopt = 0.1,
  rewire = F,
  threshold.dist = function(i) 0.25, 
  exposure.args = list(normalized = T)
  )

#getting the info and plots
diffnet_il
summary(diffnet_il)
table(diffnet_il$toa) # time of adoption
plot_diffnet(diffnet_il, slices = c(1, 2, 3, 4))
plot_adopters(diffnet_il, what = "adopt")
plot_adopters(diffnet_il, what = "cumadopt", include.legend = FALSE)
```
<br>

```{r}
#setting the seed
set.seed(1)

#running the diffusion network
diffnet_il2 <- rdiffnet(
  t = 10, 
  seed.graph = net2_matrix,
  seed.nodes = c(1, 7, 11), # chosen nodes
  seed.p.adopt = 0.1,
  rewire = F,
  threshold.dist = function(i) 0.25, 
  exposure.args = list(normalized = T)
  )

#plotting the graph
diffnet_il2
summary(diffnet_il2)
table(diffnet_il$toa) # time of adoption
plot_diffnet(diffnet_il2, slices = c(1, 2, 3, 4))
plot_adopters(diffnet_il2, what = "adopt")
plot_adopters(diffnet_il2, what = "cumadopt", include.legend = FALSE)
```
<br>

**#8. Changes in Network Structure on Diffusion**
Looking at the adopters and cumulative adopters plot, the original network had a much earlier peak in adopters, with the peak hitting a proportion of 60% around time = 2. The newer network started off very slowly (as did the original plot!) but peaked at time = 4, around 30% before plateauing and staying at 0% for the rest of the plot. 

This means that the original network had a much quicker rate of diffusion compared to the second network.

When I changed the slices on the diffusion plot (thank you for the help by the way!) I saw that the newer network had a more linear growth (hence it taking longer) whereas the original network was much exponential, which is why it maxed out at time = 3

My main takeaway is that the original plot (because it was highly structured by design) was a much more tightly connected plot compared to the new network (which had the random ties + random nodes)


<br>
<br>
<br>
<br>
<br>
<br>
<br>