---
title: 'Homework 1: Linear Regression'
author: ''
output:
  pdf_document: default
  html_document: null
editor_options:
  chunk_output_type: console
---

There are six questions (30 total points) in this assignment. The minimum increment is 1 point. Please type in your answers directly in the R Markdown file. After completion, **successfully** knitr it as an html file. Submit <span style="color:red">**both**</span>  the html file and the R Markdown file via Canvas. Please name the R Markdown file in the following format: LastName_FirstName_HW1.Rmd, e.g. Zhao_Zifeng_HW1.Rmd.


## Used Car Dataset [9 points]
The used car dataset is the one we analyzed in class. Let's read in the data stored in `UsedCar.csv`.
```{r chunk1}
total_data <- read.csv("/Users/TomTheIntern/Desktop/Mendoza/Data Folder/UsedCar.csv", header=T, stringsAsFactors=T)
```

### **Q1 [2 points]** 
Since the dependent variable `Price` is severely right-skewed, create a log-transformation of `Price` and store it as a new variable named `Log_price` within the data.frame `total_data`.
```{r Q1}

total_data$Log_price <- log(total_data$Price)

```

### **Q2 [7 points]** 
Fit a linear regression model of Log_price w.r.t. two predictors Age and Mileage, name it `lm_short`.
```{r Q2}

lm_short <- lm(Log_price ~ Age + Mileage , data = total_data)

```

#### **Q2(a) [2 points]** 
What is the $R^2$ of `lm_short`? What are the (in-sample) MAE and RMSE of `lm_short` at the *original scale*?
```{r Q2(a)}
r_2 <- summary(lm_short)$r.squared

library(forecast)

accuracy(exp(lm_short$fitted.values) - 1, total_data$Price)

MAE <- 1065.202
RMSE <- 1526.271
```
**Answer:** 

The $R^2$ value of lm_short is .8054654.

For the original scale:

The Mean Absolute Error is 1065.202

The Root Mean Square Deviation is 1526.271
<br>

#### **Q2(b) [2 point]** 
What is the estimated coefficient of `lm_short` for Age and Mileage?
```{r Q2(b)}
lm_short <- lm(Log_price ~ Age + Mileage , data = total_data)

lm_short$coefficients
```
The coefficient for Age is -1.217033e-02 and is -1.757577e-06 for Mileage

#### **Q2(c) [3 points]** 
How should we interpret the estimated coefficient of Age at the log-scale and the original scale of Price?

**Answer:** 

The change in the log price represents a percentage change for the change of price. 
Meanwhile, the change at the original price scale represents a change in the actual price.

Because both coefficients are negative, we know that as the mileage and age of a vehicle increase, the log of its price should decrease. An older car with more miles should have a lower log price than a newer car with fewer miles.

For one unit increase of Age, the log of price should go down -1.217033e-02. <br>
For one unit increase of Mileage, the log of price should go down -1.757577e-06 <br>

## Car Seat Sales Dataset [21 points]
The car seat sales dataset contains sales of child car seats at 400 different stores and the data is stored in `Carseats.csv`. It contains 9 variables, `Sales`, `CompPrice`, `Income`, `Advertising`, `Population`, `Price`, `ShelveLoc`, `Age` and `Urban`. We would like to build a linear regression model to predict `Sales` at a planned new store. The data description is as follows.

+ `Sales`: Unit sales (in thousands) at each location
+ `CompPrice`: Price charged by competitor at each location
+ `Income`: Community income level (in thousands of dollars)
+ `Advertising`: Local advertising budget for company at each location (in thousands of dollars)
+ `Population`: Population size in region (in thousands)
+ `Price`: Price company charges for car seats at each site
+ `ShelveLoc`: A factor with levels Bad, Good and Medium indicating the quality of the shelving location for the car seats at each site
+ `Age`: Average age of the local population
+ `Urban`: A factor with levels No and Yes to indicate whether the store is in an urban or rural location

####  **Q4 [2 points]**
Which variable is the dependent variable? Which predictors are categorical variables?

**Answer:**

The Dependent variable is sales, as we are trying to determine how well the new store will perform. 

The categorical predictors are ShelveLoc, which is comprised of three values and Urban, which is comprised of two. The remaining variables are all numeric.


#### **Q5 [9 points]**
Let's read in the data and perform visualization to get a better sense of the data.

#### **Q5(a) [2 points]** 
Correctly read in the data stored at `Carseats.csv`.
```{r Q5(a)}

Carseats <- read.csv("/Users/TomTheIntern/Desktop/Mendoza/Data Folder/Carseats.csv", header=T, stringsAsFactors=T)

str(Carseats)

```
<br>

####**Q5(b) [2 points]** 
Produce a scatterplot between `Sales` and `Price`. What is the general pattern from the scatterplot?
```{r Q5(b)}
library(ggplot2)
library(ggthemes)

ggplot(data = Carseats, aes(Price, Sales)) + 
  geom_point()+
  theme_stata()

```
<br>

**Answer:**
The general trend for the scatterplot is that as prices increases sales decrease. Conversely, as prices decrease, the sales increase.


#### **Q5(c) [2 points]** 
Produce a boxplot between `Sales` and `ShelveLoc`. What is the general pattern from the boxplot?
```{r Q5(c)}

ggplot(data = Carseats, aes(ShelveLoc, Sales)) + 
  geom_boxplot()+
  theme_stata()

```
<br>

**Answer:**
Sales in the Good ShelveLoc category tended to be the highest, while sales in the bad ShelveLoc were the lowest. The sales in the Medium ShelveLoc were split between the two.


#### **Q5(d) [3 points]** 
Produce a histogram of `Sales`. Is `Sales` severely right-skewed?
```{r Q5(d)}

ggplot(data = Carseats, aes(Sales)) + 
  geom_histogram(binwidth = 1)+
  theme_stata()

```
<br>

**Answer:**
Based on the plot I would saw that Sales is somewhat right skewed, but not severely right skewed.

### **Q6 [10 points]**

#### **Q6(a) [4 points]** 
Fit a linear regression model of the original scale `Sales` w.r.t. all the predictors available in the dataset, name it `lm_full`. Take a look at the summary of `lm_full`.

```{r Q6(a)}

lm_full <- lm(Sales ~ CompPrice + Income + Advertising + Population + Price + ShelveLoc + Age + Urban, 
              data = Carseats)

summary(lm_full)
```

#### **Q6(b) [2 points]** 
What is the estimated coefficient of `lm_full` for CompPrice and Price? How should we interpret the estimated coefficients?
```{r Q6(b)}

lm_full$coefficients

```

**Answer:**
When a competitors price goes up one unit, Sales for the dealer we are predicting go up. Meanwhile, if the dealer we are predicting see a raise in prices, sales for that dealer go down.

Because we know that CompPrice's coefficent is 0.0926752906 and Price's coefficent is -0.0954334977, we know that for each unit that CompPrice goes up, Sales will go up by 0.0926752906 units. 



#### **Q6(c) [2 points]** 

Which predictor(s) are not statistically significant in the model?

**Answer:** 
Population and UrbanYes


#### **Q6(d) [2 points]** 
What is the $R^2$ of `lm_full`? What are the (in-sample) MAE and RMSE of `lm_full`?
```{r Q6(d)}
#making a new model for the original scale
lm_full <- lm(Sales ~ CompPrice + Income + Advertising + Population + Price + ShelveLoc + Age + Urban, 
              data = Carseats)

summary(lm_full)

r_2_two <- summary(lm_full)$r.squared

accuracy(lm_full$fitted.values, Carseats$Sales)

MAE <- 0.805898
RMSE <- 1.006678
```

**Answer:** 
<br>
The $R^2$ of lm_full is 0.8726193
<br>
MAE: 0.805898
<br>
RMSE: 1.006678

<br>
<br>
<br>






