---
title: 'Homework 2: Data Partition and Backward Selection'
author: "Thomas Zwiller"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

There are six questions (30 total points) in this assignment. The minimum increment is 1 point. Please type in your answers directly in the R Markdown file. After completion, **successfully** knit Rr as an html file. Submit <span style="color:red">**both**</span>  the html file and the R Markdown file via Canvas. Please name the R Markdown file in the following format: LastName_FirstName_HW2.Rmd, e.g. Zhao_Zifeng_HW2.Rmd.


## Used Car Dataset [12 points]
The used car dataset is the one we analyzed in class. Let's read in the data stored in `UsedCar.csv` and further partition the data into training and test data. Note that we use the same random seed `set.seed(7)` as in class to ensure reproducibility.

```{r chunk1}
total_data <- read.csv("/Users/TomTheIntern/Desktop/Mendoza/Mod 2/Advanced Stats/Hmwk 2/UsedCar.csv", header=T, stringsAsFactors=T)
set.seed(7)
total_obs <- dim(total_data)[1]
# Data partition / Sample splitting
train_data_indices <- sample(1:total_obs, 0.8*total_obs)
train_data <- total_data[train_data_indices,]
test_data <- total_data[-train_data_indices,]
# Record the size of training data and test data
train_obs <- dim(train_data)[1]
```


### **Q1 [3 points]** Model Estimation
Instead of building linear regression models on the log-scale Price, let's build linear regression models for the original scale of Price, i.e. without log transformation to correct the right-skewness of Price.

**Q1(a) [2 points]**
Fit a linear regression model of **original scale** Price w.r.t. all 10 predictors using the **training data**, name it `lm_full`.
```{r Q1(a)}

lm_full <- lm(Price ~ . , data = train_data)
summary(lm_full)$r.squared

```

**Q1(b) [1 points]**
Check the estimated coefficient for `Mileage`, how do we interpret it?
```{r Q1(b)}

summary(lm_full)$coefficients

```

Answer: The Mileage variable has a coefficient of -1.669933e-02 meaning that for each increase of 1 mile the price goes down roughly $0.0016699. This makes sense because we know that the value of a car is negatively correlated with price: cars tend to be considered less valuable the more they are drive. 

### **Q2 [4 points]** Backward Selection with BIC

**Q2(a) [2 points]**
Perform backward selection for `lm_full` with **BIC** using the function `step()` and name the selected model `lm_bwd`. Make sure you use the correct **`k`** argument in the `step()` function.
```{r Q2(a)}

lm_bwd <- step(lm_full, direction = 'backward', k = log(nrow(train_data)))

```

**Q2(b) [2 points]**
Examine the selected model in `lm_bwd`, list all the predictors that are eliminated during the backward selection process.
```{r Q2(b)}

summary(lm_bwd)$coefficients

```

Answer: lm_bwd only uses Age, Mileage, Fuel_TypeDiesel, Fuel_TypePetrol, HP, Quarterly_Tax and Weight. 

This means that the backward selection process discarded the Doors, Automatic, CC and Metallic variables. 


### **Q3 [5 points]** Model Evaluation (Prediction)

**Q3(a) [2 points]**
Use `lm_full` and `lm_bwd` to generate predictions for Price on the test data and store the prediction in `lm_full_pred` and `lm_bwd_pred` respectively.
```{r Q3(a)}

lm_full_pred <- predict(lm_full, newdata = test_data)
lm_bwd_pred <- predict(lm_bwd, newdata = test_data)

```

**Q3(b) [2 points]**
Use the `R` package `forecast` to evaluate the prediction performance of `lm_full_pred` and `lm_bwd_pred`. What are the MAE for `lm_full` and `lm_bwd`?
```{r Q3(b)}
library(forecast)
accuracy(lm_full_pred, test_data$Price)

accuracy(lm_bwd_pred, test_data$Price)

```

Answer:

lm_full has an MAE of 1058.757 while lm_bwd has an MAE of 1069.966.

**Q3(c) [1 points]**
Recall from the in-class exercise that the MAE made by `lm_full` with log-transformation are `950.0841`. Compare with the MAE made by `lm_full` in Q3(b) without log-transformation. 
Answer:

Based on the MAE of lm_full from the in-class work and the lm_full I developed from the homework, I would say that the log transformation was able to help create a more accurate model. Because the log model has an MAE of 950.0841, it is more accurate than the non-log model, which had an MAE of 1058.757. 


## Car Seat Sales Dataset [18 points]
The car seat sales dataset is the one we analyzed in HW1. It contains sales of child car seats at 400 different stores and the data is stored in `Carseats.csv`. It contains 9 variables, `Sales`, `CompPrice`, `Income`, `Advertising`, `Population`, `Price`, `ShelveLoc`, `Age` and `Urban`. We would like to build a linear regression model to predict `Sales` at a planned new store. The data description is as follows.

+ `Sales`: Unit sales (in thousands) at each location
+ `CompPrice`: Price charged by competitor at each location
+ `Income`: Community income level (in thousands of dollars)
+ `Advertising`: Local advertising budget for company at each location (in thousands of dollars)
+ `Population`: Population size in region (in thousands)
+ `Price`: Price company charges for car seats at each site
+ `ShelveLoc`: A factor with levels Bad, Good and Medium indicating the quality of the shelving location for the car seats at each site
+ `Age`: Average age of the local population
+ `Urban`: A factor with levels No and Yes to indicate whether the store is in an urban or rural location


###  **Q4 [5 points]** Data Partition
**Q4(a) [2 points]**
Let's correctly read in the data in `Carseats.csv` and name it as `total_data`. 
```{r Q4(a)}

total_data <- read.csv( "/Users/TomTheIntern/Desktop/Mendoza/Mod 2/Advanced Stats/Hmwk 2/Carseats.csv", header = TRUE)

```

**Q4(b) [3 points]**
Let's partition the data in `total_data` into training **(80%)** and test data **(20%)** and store them as `R` objects `train_data` and `test_data` respectively. Use random seed **`set.seed(7)`**!
```{r Q4(b)}
set.seed(7)
total_obs <- dim(total_data)[1]
train_data_indices <- sample(1:total_obs, 0.8*total_obs)
train_data <- total_data[train_data_indices,]
test_data <- total_data[-train_data_indices,]
```

### **Q5 [8 points]** Model Estimation and Backward Selection

**Q5(a) [2 points]**
Fit a linear regression model of **original scale** Sales w.r.t. all 8 predictors using the **training data**, name it `lm_full`.
```{r Q5(a)}

lm_full <- lm(Sales ~ . , data = train_data)
summary(lm_full)

```

**Q5(b) [2 points]**
Perform backward selection for `lm_full` with **BIC** using the function `step()` and name the selected model `lm_bwd`. Make sure you use the correct **`k`** argument in the `step()` function.
```{r Q5(b)}

lm_bwd <- step(lm_full, direction = 'backward', k = log(nrow(train_data)))
summary(lm_bwd)

```


**Q5(c) [2 points]**
Examine the printout of the `step()` function in Q5(b), what is the first predictor removed in the backward selection?

Answer: 

Population is the first predictor removed by the backward selection process.

**Q5(d) [2 points]**
Examine the selected model in `lm_bwd`, list all the predictors that are eliminated during the backward selection process.
```{r Q5(c)}

summary(lm_bwd)

```

Answer: The backward selection process removed the population predictor and the UrbanYes predictor. 

### **Q6 [5 points]** Model Evaluation (Prediction)
**Q6(a) [2 points]**
Use `lm_full` and `lm_bwd` to generate predictions for Sales on the test data and store the prediction in `lm_full_pred` and `lm_bwd_pred` respectively.
```{r Q6(a)}

lm_full_pred <- predict(lm_full, newdata = test_data)
lm_bwd_pred <- predict(lm_bwd, newdata = test_data)


```

**Q6(b) [2 points]**
Use the `R` package `forecast` to evaluate the prediction performance of `lm_full_pred` and `lm_bwd_pred`. What are the MAE for `lm_full` and `lm_bwd`?
```{r Q6(b)}

library(forecast)
accuracy(lm_full_pred, test_data$Sales)
accuracy(lm_bwd_pred, test_data$Sales)

```

Answer: lm_full has an MAE of 0.8582053 while lm_bwd has an MAE of 0.8597975.

**Q6(c) [1 points]**
Which statistical model do you prefer, `lm_full` or `lm_bwd`? Give reasons.

Answer:

There are two ways to look at which model is better. The first is accuracy: which model has the higher $R^2$ value and which model has the lower MAE?

lm_full had the higher $R^2$ value at .8772, beating out lm_bwd narrowly (0.8768). lm_full also had a slightly lower MAE (0.8582053) thabn lm_bwd (0.8597975).

But we also want our model to have a parsimonious structure, or we want our model to use fewer predictors when possible. Predictors should only be included when they have great predictive power. So, because lm_bwd has fewer predictors, it seems like the better model. We can test this with a BIC test.

```{r}
BIC(lm_full)
BIC(lm_bwd)
```

Based on the BIC test, we can select the lm_bwd model. 







